Persistence in functional topology and a correction to a theorem of Morse

### 2021-07-29 surfaces.tex
Overleaf version to post

\section{Minimal surfaces and functional topology} \label{s:surfaces}

After having presented Morse inequalities in terms of persistence diagrams and having developed topological conditions for existence of persistence diagrams associated to filtrations, we will now connect these previous results to the work of Morse and Tompkins on minimal surfaces from \cite{Morse.1939}.
We will focus on the homological aspects of their approach.
For a general exposition we refer to \cite[Sections 4.3-5]{Bott.1980}, and for a more detailed exposition including the analytical details see \cite[Section II.6]{Struwe.1988}.

Morse and Tompkins considered the following setting introduced by Douglas.
Let $g \colon \R \to \R^n$ be a $2\pi$-periodic function representing a simple closed curve such that $g$ is differentiable with Lipschitz derivative.
Let $\widetilde{\Omega}$ be the space of continuous functions $\varphi \colon \R \to \R$ with $\varphi((t)+2\pi) = \varphi(t) + 2\pi$ for all $t$ and such that there are three distinct points $\alpha_i \in [0,2\pi)$ with $\varphi(\alpha_i)=\alpha_i$.
The \emph{Douglas functional} on $\widetilde \Omega$ associated to the curve $g$ is defined by
\begin{equation*}
A_g(\varphi)=\frac{1}{16}\int_0^{2\pi}\int_0^{2\pi}\sin\left(\frac{\alpha-\beta}{2}\right)^{-2} \! \lVert g(\varphi(\alpha))-g(\varphi(\beta)) \rVert_2^2 \ \mathrm{d}\alpha \ \mathrm{d}\beta.
\end{equation*}
Let $\Omega_g=\{\varphi\in\widetilde\Omega\mid A_g(\varphi)<\infty\}$.
Douglas proved that, for a large class of curves $g$ the set $\Omega_g$ is non-empty with compact sublevel sets of $A_g$.
Since $A_g$ is bounded below by $0$, this implies that $A_g$ has a unique global minimizer.
Douglas then used harmonic analysis to show that this minimizing critical point of $A_g$ corresponds to a solution of Plateau's Problem.

In \cite[p.~445]{Morse.1939}, Morse and Tompkins introduce a homotopical notion of critical point for any metric space $M$ and $F \colon M \to \R$.
Roughly, a point $p$ is \emph{homotopically critical} if it has no neighborhood in $M_{\leq F(p)}$ that can be mapped by a homotopy into $M_{\leq t}$ for some $t<F(p)$.

Morse and Tompkins prove that each homotopically critical point of the functional $A_g$ corresponds to a critical point of the area functional, i.e., a minimal surface, and use this correspondence to prove the following result \cite[p.~447]{Morse.1939}, also reviewed as Theorem 6.10 in \cite{Struwe.1988}.

@@

Their proof relies on Morse inequalities for homotopically critical points which they infer by relating them to cap numbers and then applying the Morse inequalities for cap numbers as we have considered them in \cref{s:inequalities}, and as Morse considered them in \cite{Morse.1940}.

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon M \to \R$ is said to be \emph{homotopically locally connected} or $\piLC$	if for any $x \in X$, $V$ a neighborhood of $x$, and any index $t > f(x)$, there is an index $s$ with $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ induces trivial maps on homotopy groups.
\end{defi}

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be \emph{locally connected} or $\LC$ if for any $x \in X$, $V$ a neighborhood of $x$, and any index $t > f(x)$, there is an index $s$ with $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.
\end{defi}

Clearly, being $\LC$ implies being $\piLC$ and, if the homology $\H$ takes finite dimensional values on one point spaces, also $\HLC$.
We will now show that not even $\LC$ is sufficient to ensure the q-tameness of compact sublevel set filtrations that are induced by non-continuous functions.
In particular, our construction will be used to invalidate Morse's claim quoted above.

@@

\begin{thm} \label{t:counterexample}
	The function $f \colon \HE \to \R$ whose value at the origin is $0$ and is $1$ everywhere else defines a compact $\LC$ sublevel set filtration that is not q-tame with respect to $\H$ if $\H_{n}(\HE)$ is not finite dimensional for some $n$.
\end{thm}

@@

	Let us now verify that the sublevel set filtration of $f$ is $\LC$, so
	consider $x \in \HE$, $V$ a neighborhood of $x$ in $\HE$ and $t > f(x)$. We need to find a neighborhood $U \subseteq V$ of $x$ and $s \in (f(x), t)$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.

@@

	For $x$ not the origin we have $f(x) = 1$ and choose $s \in (1,t)$ arbitrarily, so that $f_{\leq s} = f_{\leq t} = \HE$. 
	Note that because $x$ is not the origin, there is a unique $d$-sphere in $\HE$ that contains it.
	Clearly, we may choose $\delta > 0$ so small that $D_{\delta}(x) = \{y \in \R^{d+1} \mid \Vert x - y \Vert < \delta\} \cap \HE$ is a disk contained in this sphere and contained in $V$. 
	The disk $D_\delta(x)$ can be contracted to $\{x\}$ in $V$, so choosing $U = D_{\delta}(x)$, we obtain that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to the constant map with value $x$.

	What remains to be shown is that $f_{\leq \bullet}$ is not q-tame for $\H$.
	This follows directly from our assumption that $\H_{n}(\HE)$ is not finite dimensional for some $n$ because $f_{\leq t}$ is constant with value $\HE$ for $t \geq 1$.
\end{proof}

Both singular and \v{C}ech homology of the Hawaiian earring are infinite dimensional in some degree, so they satisfy the assumptions of \cref{t:counterexample}.
The singular case is treated in \cite{Barratt.1962}, and for \v{C}ech homology, one can use the fact that it commutes with totally ordered limits for compact Hausdorff spaces \cite[Theorems VIII.3.6.\@ and X.3.1.]{Eilenberg.1952}:
Define 
\begin{align*}
\HE_k &=
\left\{ (x_0, \dots, x_d) \in \R^{d+1} \ \middle| \ \left( x_0 - \frac{1}{k} \right)^2 + x_1^2 + \dots + x_d^2 \leq \left( \frac{1}{k} \right)^2 \right\} \\ &\cup
\bigcup_{n=1}^{k-1} \left\{ (x_0, \dots, x_d) \in \R^{d+1} \ \middle |\ \left( x_0 - \frac{1}{n} \right)^2 + x_1^2 + \dots + x_d^2 = \left( \frac{1}{n} \right)^2 \right\},
\end{align*}
i.e., the $d$-dimensional Hawaiian earring but with the $k$-th largest $d$-sphere filled.
We have $\lim_{k} \HE_{k} = \bigcap_{k} \HE_{k} = \HE$, and hence $\CH_{d}(\HE; \mathbb{F}) = \lim_{k} \CH_{d}(\HE_{k}; \mathbb{F})$, where $\CH$ denotes \v{C}ech homology.
Clearly, each $\HE_{k}$ is a CW-complex, so we can simply use cellular homology to compute
\begin{equation*}
\lim_{k}\CH_{d}(\mathbb{H}^{d}_{k}; \mathbb{F})=\lim\left(\dots\to \prod_{n=1}^2\mathbb{F}\to \prod_{n=1}^1\mathbb{F}\to \prod_{n=1}^0\mathbb{F}\right)=\prod_{n\in\mathbb{N}}\mathbb{F},
\end{equation*}
which is infinite-dimensional over $\mathbb{F}$.

\begin{cor} \label{c: counterexample}
	The function $f \colon \HE \to \R$ whose value at the origin is $0$ and is $1$ everywhere else defines a weakly $\LC$ compact sublevel set filtration that is not q-tame with respect to singular and \v{C}ech homology.
\end{cor}

The gap we have highlighted in the argument of Morse and Tompkins can be fixed by applying \cref{t:strong local connectedness implies q-tameness} because the proof given in \cite[p.464]{Morse.1939} for local connectivity of $(\Omega_g, A_g)$ actually establishes a stronger property for the sublevel set filtration of $A_g$, which we now formalize.

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be \emph{strongly} $\LC$ if for any $x \in X$, any neighborhood $V$ of $x$ and any pair of indices $f(x) < s < t$ there is a neighborhood $U \subseteq V$ of $x$ such that the map $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.
\end{defi}

Clearly, the filtration being strongly $\LC$ implies that the filtration is also strongly $\HLC$ for \v{C}ech homology with field coefficients, which is what Morse and Tompkins use.
Therefore, from \cref{t:strong local connectedness implies q-tameness} we can conclude that $(\Omega_g, A_g)$ is \mbox{q-tame} as claimed.
This implies that Morse inequalities hold, and hence so does the Morse-Tompkins' Unstable Minimal Surfaces Theorem.

We have also mentioned that Morse introduced another condition that he also called local $F$-connectivity three years earlier.
It roughly corresponds to being \emph{strongly} $\piLC$ with a certain added uniformity property.
In the original it reads:
\begin{displaycquote}[p.421-422]{Morse.1937}
	The space $M$ will be said to be locally $F$-connected for the order $n$ if corresponding to $n$, an arbitrary point $p$ on $M$, and an arbitrary positive constant $e$, there exists a positive constant $\delta$ with the following property.
	For $c \geq F(p)$ any singular $n$-sphere on $F \leq c$ (the continuous image on $F \leq c$ of an ordinary $n$-sphere) on the $\delta$-neighborhood $p_{\delta}$ of $p$ is the boundary of a singular $(n + 1)$-cell on $F \leq c + e$ and on $p_e$.
\end{displaycquote}
Morse also claims in the given reference that this condition is sufficient for q-tameness, but without providing a proof.
Whether this statement is true or not is not covered by our analysis, because the $\piLC$ and $\HLC$ conditions generally do not imply each other.
We expect the quoted claim to be true, but do not investigate it further.

>>>>

\section{Minimal surfaces and persistent homology} \label{s:surfaces}

Having established Morse inequalities in terms of persistence diagrams and topological conditions for existence of persistence diagrams associated to filtrations, we now connect these results to the work of Morse and Tompkins on minimal surfaces from \cite{Morse.1939}.
We will focus on the homological aspects of their approach.
For a general exposition we refer to \cite[Sections 4.3--5]{Bott.1980}, and for a more detailed exposition including the analytical details see \cite[Section II.6]{Struwe.1988}.

Morse and Tompkins considered the following setting introduced by Douglas.
Let $g \colon \R \to \R^n$ be a $2\pi$-periodic function representing a simple closed curve such that $g$ is differentiable with Lipschitz derivative.
Let $\widetilde{\Omega}$ be the space of continuous functions $\varphi \colon \R \to \R$ with $\varphi((t)+2\pi) = \varphi(t) + 2\pi$ for all $t$ and such that there are three distinct points $\alpha_i \in [0,2\pi)$ with $\varphi(\alpha_i)=\alpha_i$.
The \emph{Douglas functional} on $\widetilde \Omega$ associated to the curve $g$ is defined by
\begin{equation*}
A_g(\varphi) = \frac{1}{16} \int_0^{2\pi} \int_0^{2\pi} \sin \left( \frac{\alpha-\beta}{2} \right)^{-2} \! \lVert g(\varphi(\alpha)) - g(\varphi(\beta)) \rVert_2^2 \ \mathrm{d}\alpha \ \mathrm{d}\beta.
\end{equation*}
Let $\Omega_g = \{\varphi \in \widetilde\Omega \mid A_g(\varphi) < \infty\}$.
Douglas proved that, for a large class of curves $g$, the set $\Omega_g$ is non-empty and the sublevel sets of $A_g$ are compact.
Since $A_g$ is bounded below by $0$, this implies that $A_g$ has a unique global minimizer.
Using methods from harmonic analysis, Douglas then showed that this minimizing critical point of $A_g$ corresponds to a solution of Plateau's Problem.

As alluded to in \cref{r:homotopically critial points}, Morse and Tompkins consider a homotopical notion of critical point for a general function $F \colon M \to \R$ on a metric space \cite[p.~445]{Morse.1939}, see also \cite{Morse.1943}.
Roughly, a point $p$ is \emph{homotopically critical} if it has no neighborhood in $M_{\leq F(p)}$ that can be mapped by a homotopy into $M_{\leq t}$ for some $t<F(p)$.

Morse and Tompkins prove that each homotopically critical point of the functional $A_g$ corresponds to a critical point of the area functional, called a minimal surface, and use this correspondence to prove the following result \cite[p.~447]{Morse.1939}, also reviewed as Theorem II.6.10 in \cite{Struwe.1988}.

@@

Their proof relies on Morse inequalities for homotopically critical points, which they infer by relating these critical points to the cap numbers associated to the persistent \v{C}ech homology of the sublevel set filtration of the function, and then applying the Morse inequalities for cap numbers as we have considered them in \cref{s:inequalities}, and as Morse considered them in \cite{Morse.1940}.

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon M \to \R$ is said to be \emph{weakly homotopically locally connected} or \emph{weakly $\piLC$}	if for any $x \in X$, $V$ a neighborhood of $x$, and any index $t > f(x)$, there is an index $s$ with $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ induces trivial maps on homotopy groups.
\end{defi}

@@

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be \emph{weakly locally connected} or $\LC$ if for any $x \in X$, $V$ a neighborhood of $x$, and any index $t > f(x)$, there is an index $s$ with $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.
\end{defi}

Clearly, being weakly $\LC$ implies being weakly $\piLC$ and, if the homology $\H$ takes finite dimensional values on one point spaces, also weakly $\HLC$.
We will now show that not even weakly $\LC$ is sufficient to ensure the q-tameness of compact sublevel set filtrations that are induced by non-continuous functions.
In particular, our construction will be used to invalidate Morse's claim quoted above.

@@

\begin{thm} \label{t:counterexample}
	The function $f \colon \HE \to \R$ whose value at the origin is $0$ and is $1$ everywhere else defines a compact and weakly $\LC$ sublevel set filtration that is not q-tame with respect to $\H$ if $\H_{n}(\HE)$ is infinite dimensional for some $n$.
\end{thm}

@@

	In order to verify that the sublevel set filtration of $f$ is weakly $\LC$, we
	consider $x \in \HE$, $V$ a neighborhood of $x$ in $\HE$ and $t > f(x)$. We need to find a neighborhood $U \subseteq V$ of $x$ and $s \in (f(x), t)$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.

@@

	For $x$ not the origin we have $f(x) = 1$ and choose $s \in (1,t)$ arbitrarily, so that $f_{\leq s} = f_{\leq t} = \HE$. 
	Note that since $x$ is not the origin, there is a unique $d$-sphere in $\HE$ that contains $x$.
	Clearly, we may choose $\delta > 0$ so small that $D_{\delta}(x) = \{y \in \R^{d+1} \mid \Vert x - y \Vert < \delta\} \cap \HE$ is a topological disk contained in this sphere and contained in $V$. 
	The disk $D_\delta(x)$ can be contracted to $\{x\}$ in $V$, so choosing $U = D_{\delta}(x)$, we obtain that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to the constant map with value $x$.

	It remains to be shown that $f_{\leq \bullet}$ is not q-tame for $\H$.
	This follows directly from our assumption that $\H_{n}(\HE)$ is not finite dimensional for some $n$, as $f_{\leq t}$ is constant with value $\HE$ for $t \geq 1$.
\end{proof}

Both singular and \v{C}ech homology of the Hawaiian earring are infinite dimensional in some degree, so they satisfy the assumptions of \cref{t:counterexample}.
The singular case is treated in \cite{Barratt.1962}, and for \v{C}ech homology, one can use the fact that it commutes with totally ordered limits for compact Hausdorff spaces \cite[Theorems VIII.3.6.\@ and X.3.1.]{Eilenberg.1952}:
Define 
\begin{align*}
\HE_k &=
\left\{ (x_0, \dots, x_d) \in \R^{d+1} \ \middle| \ \left( x_0 - \frac{1}{k} \right)^2 + x_1^2 + \dots + x_d^2 \leq \left( \frac{1}{k} \right)^2 \right\} \\ &\, \cup
\bigcup_{n=1}^{k-1} \left\{ (x_0, \dots, x_d) \in \R^{d+1} \ \middle |\ \left( x_0 - \frac{1}{n} \right)^2 + x_1^2 + \dots + x_d^2 = \left( \frac{1}{n} \right)^2 \right\},
\end{align*}
i.e., the $d$-dimensional Hawaiian earring but with the $k$-th largest $d$-sphere filled.
We have $\lim_{k} \HE_{k} = \bigcap_{k} \HE_{k} = \HE$, and hence $\CH_{d}(\HE; \mathbb{F}) = \lim_{k} \CH_{d}(\HE_{k}; \mathbb{F})$, where $\CH$ denotes \v{C}ech homology.
Clearly, each $\HE_{k}$ is a CW-complex, so we can simply use cellular homology to compute
\begin{equation*}
\lim_{k}\CH_{d}(\mathbb{H}^{d}_{k}; \mathbb{F})=\lim\left(\dots\to \prod_{n=1}^2\mathbb{F}\to \prod_{n=1}^1\mathbb{F}\to \prod_{n=1}^0\mathbb{F}\right)=\prod_{n\in\mathbb{N}}\mathbb{F},
\end{equation*}
which is infinite-dimensional over $\mathbb{F}$.

\begin{cor} \label{c:counterexample}
	The function $f \colon \HE \to \R$ whose value at the origin is $0$ and is $1$ everywhere else defines a weakly $\LC$ compact sublevel set filtration that is not q-tame with respect to singular and \v{C}ech homology.
\end{cor}

The gap we have highlighted in the argument of Morse and Tompkins, that the sublevel set filtration of $A_g$ is not necessarily q-tame, can be fixed by applying \cref{t:local connectedness implies q-tameness}.
This is because the proof given in \cite[p.464]{Morse.1939} for the local connectivity of $(\Omega_g, A_g)$ actually establishes a stronger property, which we now formalize.

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be $\LC$ if for any $x \in X$, any neighborhood $V$ of $x$ and any pair of indices $f(x) < s < t$ there is a neighborhood $U \subseteq V$ of $x$ such that the map $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is homotopic to a constant map.
\end{defi}

Clearly, the filtration being $\LC$ implies that the filtration is also $\HLC$ for \v{C}ech homology with field coefficients, which is the homology theory Morse and Tompkins use.
Therefore, from \cref{t:local connectedness implies q-tameness} we can conclude that $(\Omega_g, A_g)$ is indeed \mbox{q-tame} as needed.
This implies that the generalized Morse inequalities hold, and hence so does the Unstable Minimal Surfaces Theorem.

We have also mentioned that Morse introduced another condition that he also called local $F$-connectivity three years earlier.
It roughly corresponds to being $\piLC$ with a certain added uniformity property.
In the original it reads:
\begin{displaycquote}[p.421--422]{Morse.1937}
	The space $M$ will be said to be locally $F$-connected for the order $n$ if corresponding to $n$, an arbitrary point $p$ on $M$, and an arbitrary positive constant $e$, there exists a positive constant $\delta$ with the following property.
	For $c \geq F(p)$ any singular $n$-sphere on $F \leq c$ (the continuous image on $F \leq c$ of an ordinary $n$-sphere) on the $\delta$-neighborhood $p_{\delta}$ of $p$ is the boundary of a singular $(n + 1)$-cell on $F \leq c + e$ and on $p_e$.
\end{displaycquote}
Morse also claims in the given reference that this condition is sufficient for q-tameness, but without providing a proof.
Whether this statement is true or not is not covered by our analysis, because the $\piLC$ and $\HLC$ conditions generally do not imply each other.
We expect the quoted claim to be true, but do not investigate it further.

### 2021-07-29 vietoris.tex
Overleaf version to post

\section{Vietoris and \v{C}ech homology} \label{s:vietoris}

@@

For greater generality, instead of metric Vietoris homology we have instead considered \v{C}ech homology in the previous section.
We now justify this choice showing that these two homology constructions agree on compact metric spaces.

First, let us recall the definition of \v{C}ech homology as presented for example by \citet[Section IX--X]{Eilenberg.1952}.
Let $X$ be a topological space and let $\Cov(X)$ be the set of all open covers of $X$ ordered by the refinement relation. 
Recall that for an open cover $\alpha \in \Cov(X)$ its \emph{nerve} $\Nrv(\alpha)$ is defined as the simplicial complex
\begin{equation*}
\Nrv(\alpha) =
\big\{ \beta \subseteq \alpha \mid \beta \text{ is finite and } \textstyle{\bigcap_{U \in \beta}} \, U \neq \emptyset \big\}.
\end{equation*}
The nerve construction defines a functor from the poset $\Cov(X)$ regarded as a category to that of simplicial complexes. 
The \emph{\v{C}ech homology with coefficients in $\mathbb{F}$} of $X$ is defined as
\begin{equation*}
\CH(X; \mathbb{F}) \ =
\lim_{\alpha \in \Cov(X)} H(\Nrv(\alpha); \mathbb{F}).
\end{equation*}
As an alternative to the nerve construction, for a cover $\alpha \in \Cov(X)$ one can define $\Vietoris(\alpha)$ as the simplicial complex
\begin{equation*}
\Vietoris (\alpha) = \left\{ \sigma \subseteq X \mid \sigma \text{ is finite and } \sigma \in U \text{ for some } U \in \alpha \right\},
\end{equation*}
which again yields a functor from $\Cov(X)$ to simplicial complexes.
This construction is dual in the sense of Dowker's Theorem \cite{Dowker.1952} to the nerve construction.
As a consequence, we have that $H (\Nrv (\alpha); \mathbb{F}) \cong H (\Vietoris (\alpha); \mathbb{F})$.
This isomorphism is natural with respect to refinement so we get an alternative description of \v{C}ech homology as 
\begin{equation*}
\CH (X; \mathbb{F}) \ \cong
\lim_{\alpha \in \Cov (X)} H (\Vietoris (\alpha); \mathbb{F}).
\end{equation*}

If $X$ is a metric space, this is still not exactly the same as the construction of metric Vietoris homology presented in \cite{Vietoris.1927} and used by Morse, which in modern notation is the limit
\begin{equation*}
\lim_{\alpha \in \Balls(X)} H (\Vietoris (\alpha); \mathbb{F}),
\end{equation*}
where 
\begin{equation*}
\Balls (X) = \left\{ ( B_{\delta} (x) )_{x \in X} \mid \delta > 0 \right\}
\subseteq \Cov (X).
\end{equation*}
However, if the metric space $X$ is compact, then $\Balls (X)$ is coinitial in $\Cov (X)$, that is to say, they both define the same limit.
Thus, we have a natural isomorphism
\begin{equation*}
\CH (X; \mathbb{F}) \ \cong \,
\lim_{\alpha \in \Balls(X)} H (\Vietoris (\alpha); \mathbb{F})
\end{equation*}
for a compact metric space $X$.
In other words, the metric Vietoris homology theory employed in Morse's setting is canonically isomorphic to the \v{C}ech homology we considered.

>>>>

\section{Vietoris and \texorpdfstring{\v{C}}{}ech homology} \label{s:vietoris}  % command preventing a hyperref warning

@@

For greater generality, instead of metric Vietoris homology in this paper we consider \v{C}ech homology instead.
We now justify this choice by showing that these two constructions for homology agree on compact metric spaces.

First, let us recall the definition of \v{C}ech homology as presented for example by \citet[Section IX--X]{Eilenberg.1952}.
Let $X$ be a topological space and let $\Cov(X)$ be the set of all open covers of $X$ ordered by the refinement relation. 
Recall that for an open cover $\alpha \in \Cov(X)$ its \emph{nerve} $\Nrv(\alpha)$ is defined as the simplicial complex
\begin{equation*}
\Nrv(\alpha) =
\big\{ \beta \subseteq \alpha \mid \beta \text{ is finite and } \textstyle{\bigcap_{U \in \beta}} \, U \neq \emptyset \big\}.
\end{equation*}
The nerve construction defines a functor from the poset $\Cov(X)$ regarded as a category to that of simplicial complexes. 
The \emph{\v{C}ech homology with coefficients in $\mathbb{F}$} of $X$ is defined as
\begin{equation*}
\CH(X; \mathbb{F}) \ =
\lim_{\alpha \in \Cov(X)} H(\Nrv(\alpha); \mathbb{F}).
\end{equation*}
As an alternative to the nerve construction, for a cover $\alpha \in \Cov(X)$ one can define $\Vietoris(\alpha)$ as the simplicial complex
\begin{equation*}
\Vietoris (\alpha) = \left\{ \sigma \subseteq X \mid \sigma \text{ is finite and } \sigma \in U \text{ for some } U \in \alpha \right\},
\end{equation*}
which again yields a functor from $\Cov(X)$ to simplicial complexes.
This construction is dual to the nerve construction in the sense of Dowker's Theorem \cite{Dowker.1952}, which asserts that the two complexes $\Nrv (\alpha)$ and $\Vietoris (\alpha)$ are homotopy equivalent.
As a consequence, we have that $H (\Nrv (\alpha); \mathbb{F}) \cong H (\Vietoris (\alpha); \mathbb{F})$.
This isomorphism is natural with respect to refinement of covers, so we get an alternative description of \v{C}ech homology as 
\begin{equation*}
\CH (X; \mathbb{F}) \ \cong
\lim_{\alpha \in \Cov (X)} H (\Vietoris (\alpha); \mathbb{F}).
\end{equation*}

If $X$ is a metric space, this is still not exactly the same as the construction of metric Vietoris homology presented in \cite{Vietoris.1927} and used by Morse, which in modern notation is the limit
\begin{equation*}
\lim_{\alpha \in \Balls(X)} H (\Vietoris (\alpha); \mathbb{F}),
\end{equation*}
where 
\begin{equation*}
\Balls (X) = \left\{ ( B_{\delta} (x) )_{x \in X} \mid \delta > 0 \right\}
\subseteq \Cov (X).
\end{equation*}
However, if the metric space $X$ is compact, then $\Balls (X)$ is coinitial in $\Cov (X)$, that is to say, they both define the same limit.
Thus, we have a natural isomorphism
\begin{equation*}
\CH (X; \mathbb{F}) \ \cong \,
\lim_{\alpha \in \Balls(X)} H (\Vietoris (\alpha); \mathbb{F})
\end{equation*}
for a compact metric space $X$.
In other words, the metric Vietoris homology theory employed in Morse's setting is canonically isomorphic to the \v{C}ech homology we consider in this paper.
